### 2020-03-26 README.Rmd

---
output: 
  github_document:
    df_print: tibble
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
Sys.setenv("CONTENTURI_HOME"= tempdir())
```

# contentid

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/contentid)](https://CRAN.R-project.org/package=contenturi)
[![R build status](https://github.com/cboettig/contentid/workflows/R-CMD-check/badge.svg)](https://github.com/cboettig/contenturi/actions)
<!-- badges: end -->

`contentid` seeks to make it easy to adopt data workflows that are based around the [hash-uri specification](https://github.com/hash-uri/hash-uri) for content-addressable storage. This very simple idea can be used to address several challenging problems in data management.

The design and implementation of `contentid` draws from [Preston](https://github.com/bio-guoda/preston), a biodiversity dataset tracker, and [Elton](https://github.com/globalbioticinteractions/elton), a command-line tool to update/clone, review and index existing species interaction datasets.

```{r eval=FALSE}
remotes::install_github("cboettig/contentid")
```

```{r}
library(contentid)
```

Consider a classic dataset, in this case, the Vostok ice core data analyzed by [Barnola et al (1987)](https://doi.org/10.1038/329408a0 "Barnola, J., Raynaud, D., Korotkevich, Y. et al. Vostok ice core provides 160,000-year record of atmospheric CO2. Nature 329, 408–414 (1987)").  For convenience, a copy of the data is included in this package.  Let's take a peak at the data now:

```{r}
vostok_co2 <- system.file("extdata", "vostok.icecore.co2", package = "contentid")
readLines(vostok_co2, n = 10)
```

We aren't concerned with parsing this data into R, we merely want an identifier we can use to refer to this data.  It has not been assigned a DOI, to my knowledge, but we can easily compute a unique identifier under the hash uri scheme:

```{r}
content_id(vostok_co2)
```

This identifier leverages the familiar URI format of other identifiers, indicating the hash of the file content as well as the algorithm (sha256) used to compute it.  Unlike other content-addressable data storage approaches (e.g., `dat`, IPFS, see <https://github.com/hash-uri/hash-uri> for details), this identifier contains no additional information beyond the content itself and is thus agnostic of specific software implementation or storage system.

Frequently we access data resources from the internet instead.  For instance, the file in question was downloaded from <http://cdiac.ornl.gov/ftp/trends/co2/vostok.icecore.co2>.  We can download this file locally and confirm it creates the same identifier (equivalent to comparing the file checksum, but using the checksum as the identifier makes it harder to overlook this step!).  `contentid` can also let us register the URL at which we found the data in a hash archive, <https://hash-archive.org>.

```{r}
# Oh no, the official version is 503 errors.  Lets use this backup
co2_url <- "https://zenodo.org/record/3678928/files/vostok.icecore.co2"
register(co2_url)
```

Note that upon registering the URL, hash-archive.org has returned the very same identifier, confirming this is the same content as our local copy. Such an archive is nothing more than a simple key-value store, indicating which URL(s) contain our content.  In this way, it can act much like a DOI re-direct, turing our identifier into resolvable, downloadable URL.  Unlike the DOI system though, this system is completely distributed -- we need not rely on any single authority to issue the identifier, and we need not resolve the identifier to a single location.  Thanks to the cryptographic properties of the hash, it is effectively impossible to generate a different identifier for the same content, or the same identifier for different content. We need not trust the authority of the registry, because we can always verify the content we receive corresponds to what we wanted by comparing hashes (or hash URIs).

```{r}
query("hash://sha256/9412325831dab22aeebdd674b6eb53ba6b7bdd04bb99a4dbb21ddff646287e37")
```

In this case, we see more than one URL has been registered containing exactly the same content. That is not surprising, as this dataset is commonly used and reproduced. Additional useful metadata, such as the file size and time-stamp the URL was registered are also reported. Notably, multiple URLs for an object could serve different purposes: we could register URLs that are more susceptible to link rot in the long term but provide higher bandwidth for downloads in the short term (e.g. S3 bucket vs scientific repository.)

It is worth noting that we still have no guarantee that these urls will not suffer link rot and cease to work -- this is not a replacement for archival storage efforts -- but the hash URI identifier allows us to avoid relying on a single storage point, and gives us a robust way to refer to an object.

## Local Registries & Local Storage

We can also store and register content locally following this same scheme.

```{r}
store(co2_url)
```

Now when we query for the hash, we see a local storage location as well:

```{r}
query("hash://sha256/9412325831dab22aeebdd674b6eb53ba6b7bdd04bb99a4dbb21ddff646287e37")
```

The `resolve()` verb (subject to change), provides a wrapper around `query` that 'resolves' the content identifier to the content.  It will check for a locally stored copy of the content first, and download (and verify) remote content otherwise:

```{r}
path <- resolve("hash://sha256/9412325831dab22aeebdd674b6eb53ba6b7bdd04bb99a4dbb21ddff646287e37")
```

Let's take a peak at the returned copy:

```{r}
readLines(path, n = 10)
```

## Programmatic long-term data access

One application of `contentid` is to support robust access of data files in R packages.  One of the most common approaches remains the distribution of data directly inside an R package, as in our example above.  This is not ideal for many reasons.  However, packages that rely on remote access of data risk link rot of those download URLs (e.g. there is not a simple and universal mechanism for resolving download URLs from DOIs, and many sources lack DOIs anyway).  Such R packages could instead register the current links in a hash archive, allowing package or analysis code to refer directly to the hash uri rather than a link that might rot.  (Clearly this resolving step still depends on the existence of some robust hash archive, a role scientific repositories could easily fill. Note the hash archive is merely a look-up table, it does not need to store any actual content).

## Comparison to DOIs and other location-based identifiers

Most identifiers, such as DOIs, EZIDs, PURLs, etc are location-based. At the heart of such identifiers is the notion of an HTTP redirect, where the DOI or other identifier resolves to a different URL.  This allows researchers to consistently resolve the same content even if the location (or at least the internet address) of the data changes.

Content URIs are completely compatible with such location-based identifiers, indeed, there is no reason why such an identifier could not be the (or one of the) download URLs resolved by the content identifier. I think it would be ideal if data repositories such as Zenodo, Dryad, or institutional repositories, which together offer some of our best capacity for archival storage, would register the locations of their content by such content hash URIs.

```{r include = FALSE}
Sys.unsetenv("CONTENTURI_HOME")

```