### 2020-04-21 paper/paper.Rmd
rename

---
documentclass: jss
author:
  - name: Carl Boettiger
    affiliation: University of California, Berkeley
    address: >
             130 Mulford Hall #3114,
             Berkeley, CA 94720-3114 USA 
    email: \email{cboettig@berkeley.edu}
    url: https://carlboettiger.info
  - name: Jorrit Poelen
    affiliation: Independently affiliated
title:
  formatted: "Content Identifiers for Reproducible Research using \\pkg{contentid}"
  # If you use tex in the formatted title, also supply version without
  plain:     "Content Identifiers for Reproducible Research"
  # For running headers, if needed
  short:     "\\pkg{contentid}: Content Identifiers"
abstract: >
          Digital Object Identifiers (DOIs) have become the gold standard for referencing persistent,
          published research artifacts such as data files.  Unfortunately, DOIs are not well-designed
          for the use case of scripted analyses common in R and other languages. Instead of writing 
          the DOI directly into the script, researchers must usually rely on URLs or local file paths
          in the script, leaving the task of obtaining external data files required to run the analysis
          in the more fragile state that DOIs were orginally introduced to solve: namely, redirecting
          to the most recent location where the data may be found.  \\pkg{contentid} implements an 
          alternative approach in which an identifier is based on the cryptographic hash of the content
          itself rather than a location-based identifier.  While many other content-based storage systems 
          exist (version-control systems such as `git` being the most widely adopted), \\pkg{contentid}
          takes a simple and elegant approach based on *unsalted* content hashes which avoids locking the
          user into a specific tool or platform, and offers a powerful mechanism for interacting with
          metadata and provenance records of research data.  I summarize the package use, discuss the broader
          landscape of content-based identifier standards and tools and design considerations involved, and 
          then close with a few examples of more complex applications of the \\pkg{contentid} approach.  
keywords:
  # at least one keyword must be supplied
  formatted: [hash, uri, doi, identifier, sha256]
  plain:     [hash, uri, doi, identifier, sha256]
preamble: >
  \usepackage{amsmath}
output: rticles::jss_article
---

# Introduction

## Code formatting

Don't use markdown, instead use the more precise latex commands:

* \proglang{Java}
* \pkg{plyr}
* \code{print("abc")}

# R code

Can be inserted in regular R markdown blocks.

```{r}
x <- 1:10
x
```