### 2020-05-08 inst/examples/benchmarking-backends.Rmd
replacing readr with base

---
output: github_document
---

Benchmark some possible backends for the registry.

```{r setup}
library(bench)
library(contentid) # remotes::install_github("cboettig/contentid", upgrade = TRUE)

knitr::opts_chunk$set(error=TRUE)
```

## Parsing directly

```{r read}
ref <- contentid::resolve("hash://sha256/598032f108d602a8ad9d1031a2bdc4bca1d5dca468981fa29592e1660c8f4883")
df <- read.delim(ref, stringsAsFactors = FALSE)

df <- dplyr::select(df, url = contentURL, id = checksum)
ex <- sample(df$url, 1e6)
```

## Base R

```{r base}

bench_time({
  id0 <- df[df$url %in% ex,]$id
})

```

## `dplyr`

```{r dplyr}
library(dplyr)
bench_time({
  id1 <- df %>% filter(url %in% ex) %>% pull(id)
})
identical(id0, id1)

## inner join is slower, and not literally the same thing
bench_time({
  id2 <- df %>% inner_join(tibble(url = ex)) %>% pull(id)
})

identical(id0, id2)

```

## `disk.frame`

A `fst`-file backed on disk storage with light-weigth dplyr semantics.

```{r disk.frame-setup, messages = FALSE}
library(disk.frame, warn.conflicts = FALSE, quietly = TRUE)
setup_disk.frame()
#options(future.globals.maxSize = Inf) # wow memory issues quickly
```

```{r disk.frame}
df_con <- disk.frame::as.disk.frame(df)
bench::bench_time({ ##
  id3 <- df_con %>% filter(url %in% ex) %>% collect() %>% pull(id)
})

identical(sort(id0), sort(id3))
```

## Thor

```{r thor}
library(thor)
## set map size to ~ 4GB to be safe
## Thor persists to local db, but whole DB must be able to fit in RAM?
dbfile <- tempfile()
env <- thor::mdb_env(dbfile, mapsize = 1048576*4e3)
env$mput(df$url, df$id)

bench_time({ # 
  id4 <- env$mget(ex) %>% as.character()
})

#identical(id0, id4)

fs::dir_info(dbfile) %>% pull(size) %>% sum()
```

## `arrow`

- Can use `.parquet` instead of `.tsv` as base file, is slightly faster than `vroom` on compressed `tsv`. (Reads in as a standard `data.frame`)
- Offers on-disk option that we can query with `dplyr` syntax (can we forgo the `dplyr` dependency though?)  
- the `dplyr` syntax is not DBI-based, and very hit-or-miss.  `filter(x %in% ...)` semantics don't always work (but don't error?).  `inner_join()` not implemented...

```{r arrow}
library(arrow)
library(dplyr, warn.conflicts = FALSE, quietly = TRUE)
pqt <- file.path(tempfile(), "arrow_dir", "df.parquet")
dir <- dirname(pqt)
dir.create(dir, recursive = TRUE)
write_parquet(df, pqt)

## parquet on disk w/ dplyr semantics
con_arw <- arrow::open_dataset(dir)
bench_time({  # 8.8s
  id5 <- con_arw %>% 
    # inner_join(tibble(url = ex), copy=TRUE) %>%  ## NO inner join
    filter(url %in% ex) %>% 
    collect() %>% pull(id)
})

identical(id0, id5)
```

```{r}
## an in memory data.frame from parquet, but reading is quite fast!
bench_time({#  1.3s
df_pqt <- read_parquet(pqt)
})

fs::file_size(pqt)
```

## `duckdb`

- On disk, standard DBI interface. 
- not on CRAN

```{r duckdb}
# install.packages("duckdb", repos=c("http://download.duckdb.org/alias/master/rstats/", "http://cran.rstudio.com"))
library(duckdb)

ddir <- fs::path(fs::path_temp(), "duckdb", "duckdb1")
fs::dir_create(fs::path_dir(ddir))
con <- DBI::dbConnect( duckdb::duckdb(), dbdir = ddir)
DBI::dbWriteTable(con, "df", df)

bench_time({
  id6 <- tbl(con, "df") %>% inner_join(tibble(url = ex), by="url", copy = TRUE) %>% pull(id)
})
identical(id0, id6)

```

```{r duckdb-filter}
bench_time({
  id6b <- tbl(con, "df") %>%  filter(url %in% ex) %>% pull(id)
})
identical(id0, id6b)
```

```{r}
dbDisconnect(con, shutdown=TRUE)
fs::dir_info(fs::path_dir(ddir)) %>% pull(size) %>% sum()
```

## `MonetDBLite`

- On disk, standard DBI interface. 
- no longer on CRAN

```{r monet}
# install.packages("MonetDBLite", repo = "https://cboettig.github.io/drat")
library(MonetDBLite)
library(DBI)
library(dplyr)

mdir <- tempfile()
con2 <- DBI::dbConnect( MonetDBLite() , dbname = mdir)
DBI::dbWriteTable(con2, "df", df)

bench_time({
  id7 <- tbl(con2, "df") %>% inner_join(tibble(url = ex), copy = TRUE) %>% pull(id)
})
identical(id0, id7)

```

```{r monetdb-filter}
### fails if ex is a big vector

# bench_time({
#  id7b <- tbl(con2, "df") %>%  filter(url %in% ex) %>% pull(id)
#  })
# identical(id0, id7b)

```

```{r}
DBI::dbDisconnect(con2, shutdown=TRUE)
rm(con2)
```

```{r}
fs::dir_info(mdir, recurse=TRUE) %>% pull(size) %>% sum()
```