### 2022-12-01 inst/examples/benchmarks.Rmd
patch

---
output: github_document
---

Benchmark some possible backends for the registry.

```{r setup}
library(bench)
library(contentid) # remotes::install_github("cboettig/contentid", upgrade = TRUE)

knitr::opts_chunk$set(error=TRUE)
```

```{r}
tsv <- "/zpool/content-store/registry.tsv"
big_registry <- vroom::vroom(tsv)

```
```{r}
devtools::load_all()
db <- init_lmdb()
```

```{r}
library(dplyr)

unique_id <- big_registry %>% count(identifier) %>% filter(n == 1) %>% left_join(big_registry) %>% select(-n)
entries <- apply(unique_id, 1, paste0, collapse = "\t")
db$mput(unique_id$identifier, entries)

unique_source <- big_registry %>% count(source) %>% filter(n == 1)  %>% left_join(big_registry) %>% select(-n)
entries <- apply(unique_source, 1, paste0, collapse = "\t")
db$mput(unique_source$source, entries)

#unique_id  %>% head(1) %>% pull(identifier) -> ex_id
#unique_source %>% filter(!is.na(identifier)) %>% pull(source) %>% head(2)
```

```{r}
bench::mark({
  history_lmdb("https://arcticdata.io/metacat/d1/mn/v2/object/17d54d9a9557dd5118edac00232cb220", db)
})
```

```{r}
bench::bench_time({
  history_tsv("https://arcticdata.io/metacat/d1/mn/v2/object/17d54d9a9557dd5118edac00232cb220", tsv)
})

```

```{r}

bench::mark({
  sources_lmdb("hash://sha256/00000359a8f49dad91e02ef42045ab843befc0478a6bcaece377fdb99fca2645", db)
})

```

```{r}
bench::bench_time({
  sources_tsv("hash://sha256/00000359a8f49dad91e02ef42045ab843befc0478a6bcaece377fdb99fca2645", tsv)
})

```

```{r}

```

```{r}
```